### 2023-03-17 paper.tex
Finalized first version

The three different models are all similar in architecture being mostly
faithful to the original transformer architecture as presented in 
\cite{vaswani_attention_2017}. However, each of them have a different approach
to using the decoder and encoder stacks of the transformer. GPT uses only
the decoder, BERT only the encoder and T5 both the encoder and decoder
\cite{radford_improving_nodate,devlin_bert_2019,raffel_exploring_2020}.

\section{Training}
All three of the models this paper discusses use similar approaches for 
training. The general idea is to first train a baseline model unsupervised 
using a large amount of unstructured text and then fine tuning the model
for each specific task after.

The objective of the unsupervised pre-training stage is to teach the model
general language understanding. The exact details between models varies but
in general, a massive number tokens are fed into to model during the course
of pre-training. Each word corresponds to a single or a few tokens depending on 
the encoding used [citation needed, same as pricing] and the number of tokens
used in training can go up to the trillions \cite{raffel_exploring_2020}. 
The model is asked to predict a missing token based on context.

@@

\begin{figure}[h]
  \centering
  \includegraphics*[scale=0.15]{img/attention_masking.png}
  \caption{
    Different attention masking patterns seen in transformer based language models.
    Dark squares indicate that the attention mechanism can see the input $i$ at time $j$.
    Light squares indicate that the access is blocked for a given pair $i$, $j$. 
    \cite{raffel_exploring_2020}
  }
\end{figure}
\begin{figure}[h]
  \centering
  \includegraphics*[scale=0.15]{img/lm_masking.png}
  \caption{
    Different masking patterns used within language models. Dark lines correspond to 
    fully-visible masking. Light gray lines indicate causal masking.
    \cite{raffel_exploring_2020}
  }
\end{figure}
\label{masking_patterns}

@@

The models this paper discusses closely follow the initial architecture
presented in the paper originally proposing the transformer 
\cite{vaswani_attention_2017}. Most of the differences between 
the models come down to training, attention layer masking and the encoder-decoder layers
\cite{raffel_exploring_2020}.

@@

As the number of parameters in the model grows past a certain threshold, it
is no longer necessary to fine-tune the model for each specific use case 
\cite{brown_language_2020}. Depending on the task at hand, it might be sufficient
to just give the model some examples of the task initially and it will (understand what you want it to do). The examples are given to
the model after training during inference and thus differs from fine-tuning. 
This is called few-shot learning. Special cases of this are one-shot learning which is
otherwise similar to few-shot learning except the number of examples is one.
Zero-shot learning refers to giving the task description using natural language,
kind of like you were explaining it to a human.

As an example \cite{brown_language_2020} used English to French translation.
The researchers gave the model between 10 to 100 pairs of examples of the
same sentence in English and French. The last pair had just an English sentence
and the model was expected to produce the corresponding sentence translated into
French.

While few-shot learning requires less data and computing power, which makes it more
accessible compared to fine-tuning a large model, at least during the writing of 
\cite{brown_language_2020} failed to produce comparable results to fine-tuning.
The number of examples given is restricted by the size of the model's context,
which for GPT-3 was 2048 tokens, around 1500 words (citation needed, pricing page has this info).

@@

Next sentence prediction takes a pair of sentences, \texttt{A} and \texttt{B}, and requests the model
to predict whether \texttt{B} follows \texttt{A} or not. In the paper, 50\% of the time \texttt{B} 
was the next sentence and 50\% of the time it was not. \cite{devlin_bert_2019}

@@

The main differentiating factor of T5 is the way it chooses to model tasks. Instead of 
tailoring each output to the specific task at hand, every task takes in a text prefix
describing the task and outputting the results as text. For example, for STS-B classification
task the model is trained to out one of a set of floating point numbers 
$x \in \{ 1+0.2k \mid k \in \{ 0, 1, ..., 20\} \}$ as a text. Many other models just 
output a number directly as a sliding scale from 1 to 5.

@@

\subsection{Multimodal Models}
Large language models have proven themselves very capable at a variety of language related
tasks. Their usefulness quickly diminishes the further we stray from their training set.
Using multimodal models (check name) however, we can widen the area of tasks the model
can complete and even achieve positive transfer, benefitting from the variety of tasks
it is trained for \cite{driess_palm-e_2023}.

@@

For example, when asking ChatGPT, which uses a slightly modified version of GPT-3 called 
GPT-3.5 [citation needed], to do a simple calculation, it answers confidently. 
With a prompt of "\texttt{what is 491*397}" we receive a response of 
"\texttt{The product of 491 and 397 is 195127.}". Unfortunately, using a calculator
leads to a different answer: 194927. Dangerously close, but not quite there.

Another example is asking the model about something they have not been trained on, for
example about future events. Prompting the model with 
"\texttt{What did Sauli Niinistö accomplish during his third term in office?}" returns
a logically perplexing response "\texttt{Sauli Niinistö is the current President of Finland, 
and he is currently serving his second consecutive term in office. 
His third term began on March 1, 2018, and it will end on February 28, 2024. ...}".
The response claims that Sauli Niinistö is currently serving his second term in office
which is correct as of 2023 and also in 2021 where the training data of GPT-3.5 ends 
[citation needed]. However, after this, the response claims that his third term began
in 2018, directly contradicting the previous statement. It is also important to mention
that there has been a two term limit since (insert year here) written into law 
[citation needed, finlex?]. Rest of the response after the first paragraph is removed.

>>>>

The three different models are all similar in architecture being mostly
faithful to the original transformer as presented in 
\cite{vaswani_attention_2017}. However, each of them have a different approach
to using the decoder and encoder stacks of the transformer. GPT uses only
the decoder, BERT only the encoder and T5 both the encoder and decoder
\cite{radford_improving_nodate,devlin_bert_2019,raffel_exploring_2020}.

\section{Training}
All three of the models this paper discusses use similar approaches for 
training. The general idea is to first train a baseline model unsupervised 
using a large amount of unstructured text and then fine tuning the model
for each specific task afterwards.

The objective of the unsupervised pre-training stage is to teach the model
general language understanding. The exact details between models varies but
in general, a massive number tokens are fed into to model during the course
of pre-training. Each word corresponds to a single or a few tokens depending on 
the encoding used \cite{noauthor_pricing_nodate} and the number of tokens
used in training can go up to the trillions \cite{raffel_exploring_2020}. 
The model is asked to predict a missing token based on context.

@@

\begin{figure}[h]
  \centering
  \includegraphics*[scale=0.15]{img/attention_masking.png}
  \caption{
    Different attention masking patterns seen in transformer based language models.
    Dark squares indicate that the attention mechanism can see the input $i$ at time $j$.
    Light squares indicate that the access is blocked for a given pair $i$, $j$. 
    \cite{raffel_exploring_2020}
  }
\end{figure}
\begin{figure}[h]
  \centering
  \includegraphics*[scale=0.15]{img/lm_masking.png}
  \caption{
    Different masking patterns used within language models. Dark lines correspond to 
    fully-visible masking. Light gray lines indicate causal masking.
    \cite{raffel_exploring_2020}
  }
  \label{masking_patterns}
\end{figure}

@@

The models this paper discusses closely follow the initial architecture
presented in the paper originally proposing the transformer 
\cite{vaswani_attention_2017}. Most of the differences between 
the models come down to training, attention layer masking and the use of encoder-decoder layers
\cite{raffel_exploring_2020}.

@@

As the number of parameters in the model grows past a certain threshold, it
is no longer necessary to fine-tune the model for each specific use case 
\cite{brown_language_2020}. Depending on the task at hand, it might be sufficient
to just give the model some examples of the task initially and it will pick up on the task pattern. The examples are given to
the model after training during inference and thus differs from fine-tuning. 
This is called few-shot learning. Special cases of this are one-shot learning which is
otherwise similar to few-shot learning except the number of examples is one.
Zero-shot learning refers to giving the task description using natural language,
kind of like you were explaining it to a human.

As an example, \cite{brown_language_2020} used English to French translation.
The researchers gave the model between 10 to 100 pairs of examples of the
same sentence in English and French. The last pair had just an English sentence
and the model was expected to produce the corresponding sentence translated into
French.

While few-shot learning requires less data and computing power, which makes it more
accessible compared to fine-tuning a large model, at least during the writing of 
\cite{brown_language_2020} failed to produce comparable results to fine-tuning.
The number of examples given is restricted by the size of the model's context,
which for GPT-3 was 2048 tokens, around 1500 words \cite{noauthor_pricing_nodate}.

@@

Next sentence prediction takes a pair of sentences, \texttt{A} and \texttt{B}, and asks the model
to predict whether \texttt{B} follows \texttt{A} or not. In the paper, 50\% of the time \texttt{B} 
was the next sentence and 50\% of the time it was not. \cite{devlin_bert_2019}

@@

The main differentiating factor of T5 is the way it chooses to model tasks. Instead of 
tailoring each output to the specific task at hand, every task takes in a text prefix
describing the task and outputting the results as text. For example, for STS-B classification
task the model is trained to output one of a set of floating point numbers 
$x \in \{ 1+0.2k \mid k \in \{ 0, 1, ..., 20\} \}$ as a text. Many other models just 
output a number directly as a sliding scale from 1 to 5.

@@

\subsection{Multimodal Models}
Large language models have proven themselves very capable at a variety of language related
tasks. Their usefulness quickly diminishes the further we stray from their training set.
Using multimodal models however, we can widen the area of tasks the model
can complete and even achieve positive transfer, benefitting from the variety of tasks
it is trained for \cite{driess_palm-e_2023}.

@@

For example, when asking ChatGPT, which uses a slightly modified version of GPT-3 called 
GPT-3.5 \cite{noauthor_introducing_nodate-1}, to do a simple calculation, it answers confidently. 
With a prompt of "\texttt{what is 491*397}" we receive a response of 
"\texttt{The product of 491 and 397 is 195127.}". Unfortunately, using a calculator
leads to a different answer: 194927. Dangerously close, but not quite there.

Another example is asking the model about something they have not been trained on, for
example about future events. Prompting the model with 
"\texttt{What did Sauli Niinistö accomplish during his third term in office?}" returns
a logically perplexing response "\texttt{Sauli Niinistö is the current President of Finland, 
and he is currently serving his second consecutive term in office. 
His third term began on March 1, 2018, and it will end on February 28, 2024. ...}".
The response claims that Sauli Niinistö is currently serving his second term in office
which is correct as of 2023 and also in 2021 where the training data of GPT-3.5 ends. 
However, after this, the response claims that his third term began
in 2018, directly contradicting the previous statement. It is also important to mention
that there has been a two term limit since the 1990s written into law 
\cite{oy_finlex_nodate}. Rest of the response after the first paragraph is removed.
